EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-06-01 03_edipack2.tex
Full coloring of the iRDM/normal formulae

A pure quantum state $\ket{\psi}$ belonging to a unique symmetry 
sector $\SS_{\vec{Q}}$ can be represented in the Fock basis as:
$
\ket{\psi} = \sum_I a_I \ket{I},
$
where the corresponding pure density matrix $\rho$ is given by:
\begin{equation}\label{DefRho}
\rho = \ket{\psi}\bra{\psi} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
a^*_J a_I \ket{I}\bra{J} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
\rho_{IJ} \ket{I}\bra{J}.
\end{equation}
The iRDM is obtained by tracing out the bath degrees of freedom:
\begin{equation}
\rho^\mathrm{imp} = \Tr_\mathrm{bath}(\rho).
\end{equation}
However, the memory footprint of $\rho$, as well as the CPU footprint 
of the summations involved in this trace scales exponentially 
with the system size, i.e.~$4^{N_\mathrm{s}}$, quickly becoming prohibitive.

To overcome the issue, we implement a fast algorithm that exploits the 
block structure of the Fock space defined by the symmetry sectors and
the subsequent map sparsity, to perform the trace {\it on-the-fly}. 
This approach significantly reduces the computational cost by limiting 
the summation to the sector size $D_\SS$ of the symmetry sector 
$\SS$, to which the ground state (or any eigenstate) belongs, and avoiding
the storage of the large matrix associated with $\rho$.

To efficiently manage this decomposition, we employ a 
{\tt sparse\_map} ${\cal P}$, as detailed in \secu{CodeSparseMap}. 
This structure stores, for each impurity configuration 
(the {\it key}), the 
corresponding bath configurations 
(the {\it values}) that 
satisfy the symmetry constraints over the sector quantum numbers
imposed by the {\tt ed\_mode} variable.
We denote the number of (spin-dependent) keys and values respectively
as ${\cal D}_{i\s}$ and ${\cal D}_{b\s}$.
The fast summation algorithm for evaluating the iRDM differs 
significantly between the {\bf normal} and {\bf superc/nonsu2} modes, 
as outlined in the sections below.

\paragraph{{\bf normal}.}
In this case, we take advantage of spin-resolved sector decomposition, 
as discussed in detail in 
\secu{sSecQNs}. 
The global bitset can be split into spin-resolved 
impurity and bath components as
\begin{equation}
\ket{\vec{n}} =
\ket{\vec{n}_\up}\otimes \ket{\vec{n}_\dw} =
\ket{\vec{i}_\up}\ket{\vec{b}_\up}\otimes \ket{\vec{i}_\dw}\ket{\vec{b}_\dw}.
\label{eq:decomposition_for_rdm_normal}
\end{equation}
Here, creation (destruction) operators  $c^\dagger_{p\sigma}$
($c^\dagger_{p\sigma}$) at position $p$ and for a given spin $\sigma$
act only on the corresponding 
spin subspace, i.e.~they commute with the opposite spin 
subspace $\ket{\vec{n}_{\bar{\s}}}$ without introducing any 
fermionic sign. In this basis \equ{DefRho} takes the form:
\begin{equation}\label{RhoNormal}
\rho = 
    \sum_{i_\up=1}^{D_{i_\up}}\!
    \sum_{p_\up=1}^{D_{p_\up}}\!
    \sum_{i_\dw=1}^{D_{i_\dw}}\!
    \sum_{p_\dw=1}^{D_{p_\dw}}\!
    \sum_{j_\up=1}^{D_{j_\up}}\!
    \sum_{q_\up=1}^{D_{q_\up}}\!
    \sum_{j_\dw=1}^{D_{j_\dw}}\!
    \sum_{q_\dw=1}^{D_{q_\dw}}\!
    a_{i_\up p_\up i_\dw p_\dw}a^*_{j_\up q_\up j_\dw q_\dw}
    % \left[
    \ket{\vec{i}_\up}\ket{\vec{p}_\up}\!\otimes\!\ket{\vec{i}_\dw}\ket{\vec{p}_\dw}
    %\right]
    % \left[
    \bra{\vec{q}_\dw}\bra{\vec{j}_\dw}\!\otimes\!\bra{\vec{q}_\up}\bra{\vec{j}_\up}
    %\right].   
\end{equation}
where the coefficients $a_I = a_{i_\up b_\up i_\dw b_\dw}$ are the
expansion coefficients of the state on the Fock basis
$\ket{\psi} = \sum_I a_I \ket{\vec{i}_\up}\ket{\vec{b}_\up}\otimes \ket{\vec{i}_\dw}\ket{\vec{b}_\dw}$.

Before proceeding further we show that in the {\bf normal} mode the iRDM is restricted to 
{spin-diagonal} blocks. This follows from the observation 
that the conserved quantum numbers are 
{\it additive} quantities, i.e. 
$N_\s = N_\s^\mathrm{imp} + N_\s^\mathrm{bath}$,
so that the corresponding generator of the symmetry group factorizes as
$U_{N_\s} = U_{N_\s^\mathrm{imp}} \otimes U_{N_\s^\mathrm{bath}}$. 
Since the iRDM acts only on the impurity Fock space, this implies:
\begin{equation}
\begin{aligned}
\rho^\mathrm{imp} &= U_{N_\s}^\dagger \rho^\mathrm{imp} U_{N_\s} = 
U_{N_\s^\mathrm{bath}}^\dagger \otimes U_{N_\s^\mathrm{imp}}^\dagger 
\rho^\mathrm{imp} 
U_{N_\s^\mathrm{imp}} \otimes U_{N_\s^\mathrm{bath}} \\
&= 
U_{N_\s^\mathrm{imp}}^\dagger \rho^\mathrm{imp} U_{N_\s^\mathrm{imp}},
\end{aligned}
\end{equation}
showing that $U_{N_\s^\mathrm{imp}}$ defines a symmetry of 
$\rho^\mathrm{imp}$, which is therefore block diagonal with respect to 
the impurity quantum numbers $N_\s^\mathrm{imp}$.

Using \equ{RhoNormal} the iRDM takes the form:
\begin{equation}
  \label{iRDMnormal}
  \begin{aligned}
  \rho^\mathrm{imp} &= \Tr_{{\color{red}b_\up b_\dw}}{\rho} = 
  \sum_{{\color{Red} b_\up=1}}^{{\color{red} D_{b_\up}}}
  \sum_{{\color{Red} b_\dw=1}}^{{\color{red} D_{b_\dw}}}
  \bra{b_\up}\otimes\bra{b_\dw}
    \rho
    \ket{b_\dw}\otimes\ket{b_\up}    \cr
    %
    &=
    \sum_{{\color{Red} b_\up=1}}^{{\color{red} D_{b_\up}}}
    \sum_{i_\up=1}^{D_{i_\up}}
    \sum_{p_\up=1}^{D_{p_\up}}
    \sum_{j_\up=1}^{D_{j_\up}} 
    \sum_{q_\up=1}^{D_{q_\up}}
    \Bigg(    
    \sum_{{\color{Red} b_\dw=1}}^{{\color{red} D_{b_\dw}}}
    \sum_{i_\dw=1}^{D_{i_\dw}}
    \sum_{p_\dw=1}^{D_{p_\dw}}
    \sum_{j_\dw=1}^{D_{j_\dw}} 
    \sum_{q_\dw=1}^{D_{q_\dw}}
    C_{i,p}\,C_{j,q}\,a_{i_\up p_\up i_\dw p_\dw} a^*_{j_\up q_\up j_\dw q_\dw} \cr
    %
    &\qquad    
    \langle b_\up| p_\up\rangle| i_\up\rangle\langle j_\up |\langle q_\up| b_\up\rangle   \otimes \langle b_\dw| p_\dw \rangle|i_\dw\rangle \langle j_\dw  |\langle q_\dw| b_\dw\rangle \Bigg)
    \cr
&=
    \sum_{i_\up=1}^{D_{i_\up}}
    \sum_{j_\up=1}^{D_{j_\up}}
    \sum_{i_\dw=1}^{D_{i_\dw}}
    \sum_{j_\dw=1}^{D_{j_\dw}}
    \underbrace{\left(
    \sum_{{\color{Red} b_\up=1}}^{{\color{red} D_{b_\up}}}
    \sum_{{\color{Red} b_\dw=1}}^{{\color{red} D_{b_\dw}}}
    a_{i_\up b_\up i_\dw b_\dw} a^*_{j_\up b_\up j_\dw b_\dw}
    \right)}_{\rho^\mathrm{imp}_{i_\up j_\up} \otimes\, \rho^\mathrm{imp}_{i_\dw j_\dw} }
    \ket{i_\up}\bra{j_\up}\otimes \ket{i_\dw}\bra{j_\dw},
  \end{aligned}
\end{equation}

Notably, the factors $C_{i,b}$, which account for the potential 
fermionic sign from reordering bath and impurity components, are 
trivially 1 due to the absence of cross-spin interference in this 
mode, as enforced by the Kronecker deltas for bath and impurity 
indices.

>>>>

A pure quantum state $\ket{\psi}$ belonging to a unique symmetry 
sector $\SS_{\vec{Q}}$ can be represented in the Fock basis as:
$
\ket{\psi} = \sum_I a_I \ket{I},
$
where the corresponding pure density matrix $\rho$ is given by:
\begin{equation}\label{DefRho}
\rho = \ket{\psi}\bra{\psi} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
a^*_J a_I \ket{I}\bra{J} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} 
\rho_{IJ} \ket{I}\bra{J}.
\end{equation}
The iRDM is obtained by tracing out the bath degrees of freedom:
\begin{equation}
\rho^\mathrm{imp} = \Tr_\mathrm{bath}(\rho).
\end{equation}
However, the memory footprint of $\rho$, as well as the CPU footprint 
of the summations involved in the trace, scale exponentially 
with the system size, i.e.~as $4^{N_\mathrm{s}}$, quickly becoming prohibitive.
To overcome the issue, we implement a fast algorithm that exploits the 
block structure of the Fock space defined by the symmetry sectors and
the subsequent map sparsity, to perform the trace {\it on-the-fly}. 
This approach significantly reduces the computational cost by limiting 
the summation to the size $D_\SS$ of the symmetry sector 
$\SS$, to which the ground state (or any eigenstate) belongs, and avoiding
the storage of the large matrix associated with $\rho$.
We employ a 
{\tt sparse\_map} ${\cal P}$, as detailed in \secu{CodeSparseMap}. 
This structure stores, for each impurity configuration 
(the {\it key}), the 
corresponding bath configurations 
(the {\it values}) that 
satisfy the symmetry constraints over the sector quantum numbers
imposed by the {\tt ed\_mode} variable.
We denote the number of (spin-dependent) keys and values respectively
as ${D}_{i\s}$ and ${D}_{b\s}$.
The fast summation algorithm for evaluating the iRDM differs 
significantly between the {\bf normal} and {\bf superc/nonsu2} modes, 
as outlined in the sections below.

\paragraph{{\bf normal}.}
In this case, we take advantage of spin-resolved sector decomposition, 
as discussed in detail in 
\secu{sSecQNs}. 
The global bitset can be split into spin-resolved 
impurity and bath components as
\begin{equation}
\ket{\vec{n}} =
\ket{{\color{xkcdRed}\vec{n}_\up}}\otimes \ket{{\color{xkcdBlue}\vec{n}_\dw}} =
\ket{{\color{xkcdRaspberry}\vec{i}_\up}}\ket{{\color{xkcdPumpkin}\vec{b}_\up}}\otimes \ket{{\color{xkcdPurple}\vec{i}_\dw}}\ket{{\color{xkcdTeal}\vec{b}_\dw}}.
\label{eq:decomposition_for_rdm_normal}
\end{equation}
Here, creation (destruction) operators  $c^\dagger_{p\sigma}$
($c^\dagger_{p\sigma}$) at position $p$ and for a given spin $\sigma$
act only on the corresponding 
spin subspace, i.e.~they commute with the opposite spin 
subspace $\ket{\vec{n}_{\bar{\s}}}$ without introducing any 
fermionic sign. In this basis \equ{DefRho} takes the form:
\begin{equation}\label{RhoNormal}
\rho = 
    \sum_{{\color{xkcdRaspberry}i_\up}=1}^{D_{i_\up}}\!
    \sum_{{\color{xkcdPumpkin}p_\up}=1}^{D_{p_\up}}\!
    \sum_{{\color{xkcdPurple}i_\dw}=1}^{D_{i_\dw}}\!
    \sum_{{\color{xkcdTeal}p_\dw}=1}^{D_{p_\dw}}\!
    \sum_{{\color{xkcdRaspberry}j_\up}=1}^{D_{j_\up}}\!
    \sum_{{\color{xkcdPumpkin}q_\up}=1}^{D_{q_\up}}\!
    \sum_{{\color{xkcdPurple}j_\dw}=1}^{D_{j_\dw}}\!
    \sum_{{\color{xkcdTeal}q_\dw}=1}^{D_{q_\dw}}\!
    a_{{\color{xkcdRaspberry}i_\up} {\color{xkcdPumpkin}p_\up} {\color{xkcdPurple}i_\dw} {\color{xkcdTeal}p_\dw}}a^*_{{\color{xkcdRaspberry}j_\up} {\color{xkcdPumpkin}q_\up} {\color{xkcdPurple}j_\dw} {\color{xkcdTeal}q_\dw}}
    % \left[
    \ket{{\color{xkcdRaspberry}\vec{i}_\up}}\ket{{\color{xkcdPumpkin}\vec{p}_\up}}\!\otimes\!\ket{{\color{xkcdPurple}\vec{i}_\dw}}\ket{{\color{xkcdTeal}\vec{p}_\dw}}
    %\right]
    % \left[
    \bra{{\color{xkcdTeal}\vec{q}_\dw}}\bra{{\color{xkcdPurple}\vec{j}_\dw}}\!\otimes\!\bra{{\color{xkcdPumpkin}\vec{q}_\up}}\bra{{\color{xkcdRaspberry}\vec{j}_\up}}
    %\right].   
\end{equation}
where the coefficients $a_I = a_{{\color{xkcdRaspberry}i_\up} {\color{xkcdPumpkin}b_\up} {\color{xkcdPurple}i_\dw} {\color{xkcdTeal}b_\dw}}$ are the
expansion coefficients of the state on the Fock basis
$\ket{\psi} = \sum_I a_I \ket{{\color{xkcdRaspberry}{i}_\up}}\ket{{\color{xkcdPumpkin}{b}_\up}}\otimes \ket{{\color{xkcdPurple}{i}_\dw}}\ket{{\color{xkcdTeal}{b}_\dw}}$.
Before proceeding further, we show that in the {\bf normal} mode the iRDM is restricted to 
{spin-diagonal} blocks. This follows from the observation 
that the conserved quantum numbers are 
{\it additive} quantities, i.e. 
$N_\s = N_\s^\mathrm{imp} + N_\s^\mathrm{bath}$,
so that the corresponding generator of the symmetry group factorizes as
$U_{N_\s} = U_{N_\s^\mathrm{imp}} \otimes U_{N_\s^\mathrm{bath}}$. 
Since the iRDM acts only on the impurity Fock space, this implies:
\begin{equation}
\begin{aligned}
\rho^\mathrm{imp} &= U_{N_\s}^\dagger \rho^\mathrm{imp} U_{N_\s} = 
U_{N_\s^\mathrm{bath}}^\dagger \otimes U_{N_\s^\mathrm{imp}}^\dagger 
\rho^\mathrm{imp} 
U_{N_\s^\mathrm{imp}} \otimes U_{N_\s^\mathrm{bath}} \\
&= 
U_{N_\s^\mathrm{imp}}^\dagger \rho^\mathrm{imp} U_{N_\s^\mathrm{imp}},
\end{aligned}
\end{equation}
showing that $U_{N_\s^\mathrm{imp}}$ defines a symmetry of 
$\rho^\mathrm{imp}$, which is therefore block diagonal with respect to 
the impurity quantum numbers $N_\s^\mathrm{imp}$.

Using \equ{RhoNormal} the iRDM takes the form:
\begin{equation}
  \label{iRDMnormal}
  \begin{aligned}
  \rho^\mathrm{imp} &= \Tr_{{\color{xkcdCoral}b_\up} {\color{xkcdAzure}b_\dw}}{\rho} = 
  \sum_{{\color{xkcdCoral} b_\up}=1}^{D_{b_\up}}
  \sum_{{\color{xkcdAzure} b_\dw}=1}^{D_{b_\dw}}
  \bra{{\color{xkcdCoral}b_\up}}\otimes\bra{{\color{xkcdAzure}b_\dw}}
    \rho
    \ket{{\color{xkcdAzure}b_\dw}}\otimes\ket{\color{xkcdCoral}{b_\up}}    \cr
    %
    &=
    \sum_{{\color{xkcdCoral} b_\up}=1}^{D_{b_\up}}
    \sum_{{\color{xkcdRaspberry}i_\up}=1}^{D_{i_\up}}\!
    \sum_{{\color{xkcdPumpkin}p_\up}=1}^{D_{p_\up}}\!
    \sum_{{\color{xkcdRaspberry}j_\up}=1}^{D_{j_\up}}\!
    \sum_{{\color{xkcdPumpkin}q_\up}=1}^{D_{q_\up}}\!
    \Bigg(    
    \sum_{{\color{xkcdAzure} b_\dw}=1}^{D_{b_\dw}}
    \sum_{{\color{xkcdPurple}i_\dw}=1}^{D_{i_\dw}}\!
    \sum_{{\color{xkcdTeal}p_\dw}=1}^{D_{p_\dw}}\!
    \sum_{{\color{xkcdPurple}j_\dw}=1}^{D_{j_\dw}}\!
    \sum_{{\color{xkcdTeal}q_\dw}=1}^{D_{q_\dw}}\!
    C_{i,p}\,C_{j,q}\,a_{{\color{xkcdRaspberry}i_\up} {\color{xkcdPumpkin}p_\up} {\color{xkcdPurple}i_\dw} {\color{xkcdTeal}p_\dw}}a^*_{{\color{xkcdRaspberry}j_\up} {\color{xkcdPumpkin}q_\up} {\color{xkcdPurple}j_\dw} {\color{xkcdTeal}q_\dw}} \cr
    %
    &\qquad    
    \langle {\color{xkcdCoral}b_\up}| {\color{xkcdPumpkin}p_\up}\rangle| {\color{xkcdRaspberry}i_\up}\rangle\langle {\color{xkcdRaspberry}j_\up} |\langle {\color{xkcdPumpkin}q_\up}| {\color{xkcdCoral}b_\up}\rangle   \otimes \langle {\color{xkcdAzure}b_\dw}| {\color{xkcdTeal}p_\dw} \rangle|{\color{xkcdPurple}i_\dw}\rangle \langle {\color{xkcdPurple}j_\dw}  |\langle {\color{xkcdTeal}q_\dw}| {\color{xkcdAzure}b_\dw}\rangle \Bigg)
    \cr
&=
    \sum_{i_\up=1}^{D_{i_\up}}
    \sum_{j_\up=1}^{D_{j_\up}}
    \sum_{i_\dw=1}^{D_{i_\dw}}
    \sum_{j_\dw=1}^{D_{j_\dw}}
    \underbrace{\left(
    \sum_{{\color{xkcdCoral} b_\up}=1}^{D_{b_\up}}
    \sum_{{\color{xkcdAzure} b_\dw}=1}^{D_{b_\dw}}
    a_{{\color{xkcdRaspberry}i_\up} {\color{xkcdCoral}b_\up} {\color{xkcdPurple}i_\dw} {\color{xkcdAzure}b_\dw}}a^*_{{\color{xkcdRaspberry}j_\up} {\color{xkcdCoral}b_\up} {\color{xkcdPurple}j_\dw} {\color{xkcdAzure}b_\dw}}
    \right)}_{\rho^\mathrm{imp}_{{\color{xkcdRaspberry}i_\up j_\up}} \otimes\, \rho^\mathrm{imp}_{{\color{xkcdPurple}i_\dw j_\dw}} }
    \ket{{\color{xkcdRaspberry}i_\up}}\bra{{\color{xkcdRaspberry}j_\up}}\otimes \ket{{\color{xkcdPurple}i_\dw}}\bra{{\color{xkcdPurple}j_\dw}},
  \end{aligned}
\end{equation}

Notably, the factors $C_{i,b}$, which account for the 
fermionic sign associated to swapping bath and impurity components, are 
trivially 1 due to the absence of cross-spin interference in this 
mode, as enforced by the Kronecker deltas for bath and impurity 
indices.

### 2025-06-01 05_examples.tex
Full coloring of the iRDM/normal formulae

The code implementation follows the same guidelines discussed above for the other examples. The user is required to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone. 
This is employed to evaluate the local interacting Green's function:
$$
G_\mathrm{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z)
\right]^{-1}, 
$$
entering in self-consistency conditions $\GG^{-1}=G_\mathrm{loc}^{-1}+\Sigma$. 
Here  $\z\in\CCC$, $N_k$ is the number of $k$-points   and $\Sigma(z)$ is the self-energy matrix  obtained from the DMFT solution of the problem.  
Moreover, the $H(k)$ is used to  construct the {\it renormalized} topological Hamiltonian \cite{Wang2010PRL,Wang2012PRX,Blason2023PRB}
$$
H_\mathrm{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant matrix. The matrix $H_\mathrm{top}$ describes the
low-energy properties of the many-body solution andcan be used to determine the characterize the topological nature of the interacting system  \cite{Gurarie2011PRB,Wang2012PRX,Wagner2023NC,Blason2023PRB,Bau2024PRB}.

@@

To complete the overview of \NAME features, we consider another
model of interacting quantum spin-Hall insulator defined on the honeycomb lattice, with two inequivalent sites in the unit cell. Neglecting the Rashba spin-orbit interaction and discarding any ionic character of the unit cell, the noninteracting Kane-Mele Hamiltonian \cite{Kane2005PRLa,Kane2005PRL} can be written in momentum space as
\begin{align*}
    H_\mathrm{KM} = \sum_{{k}} \psi^\dagger_{{k}} H({k}) \psi_{{k}},
\end{align*}
with $\psi_{{k}} = [c_{{k},\mathrm{A},\up},\,c_{{k},\mathrm{B},\up},\,c_{{k},\mathrm{A},\dw},\,c_{{k},\mathrm{B},\dw}]$ and
    \begin{align}
    H({k}) 
    &= x_{{k}}\Gamma_{01} + y_{{k}}\Gamma_{02} + \delta_{{k}}\Gamma_{33}, \label{eq:Hk_kanemele}\\[1mm]
   x_{{k}}&=t\bigl[\cos({k}\cdot{u})+\cos({k}\cdot{v})+1\bigr],\nonumber\\
   y_{{k}}&=t\bigl[\sin({k}\cdot{u})+\sin({k}\cdot{v})\bigr],\nonumber \\
   \delta_{{k}}&=2\lambda_\mathrm{so}\bigl[\sin({k}\cdot{u})-\sin({k}\cdot{v}) - \sin({k}\cdot{u} -{k}\cdot{v})\bigr], \nonumber
\end{align}
where ${u} = (3a/2, \sqrt{3}a/2)$, and ${v} = (3a/2, -\sqrt{3}a/2)$ 
are suitable basis vectors for the honeycomb lattice, $t$ is the
nearest-neighbor hopping amplitude and $\lambda_\mathrm{so}$ is 
the amplitude of a complex next-nearest-neighbor hopping,
with a spin-dependent $\pm\tfrac{\pi}{2}$ chiral phase arising from spin-orbit coupling, which is well-known to open a topological gap at the Fermi level \cite{Kane2005PRLa,Kane2005PRL}.
The spinorial $4\times4$ matrices $\Gamma_{a\ell} = \sigma_a\otimes\tau_\ell$ are 
defined in terms of Pauli matrices $\sigma_a$, $\tau_\ell$ referred, respectively 
to spin and sublattice degrees of freedom.
% , with the convention that 
% $\sigma_0$ and $\tau_0$ are $2\times2$ identity matrices.
To investigate the effects of electron-electron repulsion we consider a local Hubbard term \cite{Hohenadler2013ROPIP,Rachel2018ROPIP}:
\begin{equation}
    H = H_\mathrm{KM} 
    + U\sum_{i}\Biggl[
                        \left(n_{i,\up} - \frac{1}{2}\right)
                        \left(n_{i,\dw} - \frac{1}{2}\right)
                      \Biggr],
    \label{eq:KMHmodel}
\end{equation}
where $n_{i,\sigma}=c^\dagger_{i,\sigma}c_{i,\sigma}$ are the 
local spin-density operators. The interaction is written in an explicitly 
particle-hole symmetric form so that the model is at half-filling for zero
chemical potential.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/flakes_edipack.pdf}\\[1mm]
    (c) \hspace{10cm} (d) \\
    \hspace{1.2cm}
    \includegraphics[width=0.33\linewidth]{figures/3NflakeU4.5_x.pdf} \hspace{1.2cm}
    \includegraphics[width=0.33\linewidth]{figures/3NflakeU6_x.pdf} \vspace{3mm}
    \caption{Kane-Mele-Hubbard model at the nanoscale.
    In (a) the local in-plane
    magnetization $\ibra S_x \iket = \ibra n_{i,\rightarrow}-n_{i,\leftarrow}\iket$. In
    (b) the local correlation as measured by the mutual 
    information between local in-plane spin eigenstates
    $I(\rightarrow:\leftarrow)$ \cite{BellomiaPhD,BellomiaKMH,Bellomia_intracorr}. 
    Both the (a) and (b) panels mark the edge and "bulk" 
    sites in lighter and darker colors, respectively.
    In (c) and (d) a visual representation of the given 
    nano-system at different values of $U/t$, as marked 
    respectively by the orange and red vertical lines 
    in panels (a) and (b). The area of the 
    circles is proportional to $I(\rightarrow:\leftarrow)$, 
    the length of the arrows to $\ibra S_x \iket$.
    All data are taken at $\lambda_\mathrm{so}=0.3t$ and $T=0$.}
    \label{fig:KMflake}
\end{figure}

\subsubsection{Kane-Mele-Hubbard model at the nanoscale}
The capabilities of the EDIpack2ineq layer go well beyond the
simple case of a two-site unit cell. Bulding on the previous example, we can readily adapt the calculation to deal with a nanoscopic~\cite{Amaricci2014PRA} system with $N_\mathrm{ineq}$ inequivalent atoms, such as hexagonal flakes
\cite{Valli2016PRB,Valli2018NL}.
We consider a 
$N_\mathrm{ineq}\times N_\mathrm{ineq}$ local Hamiltonian 
$H_\mathrm{loc}$ with open boundary conditions, and implementing
the self-consistent equations with a trivial single momentum 
$k=0$. In this configuration the system develops an inhomogeneous magnetization, so that in general no particular lattice symmetry group, e.g. N\'eel, can be exploited. 
% Some lowered point symmetry can be found, but a 
% safer approach of solving all the $N_\mathrm{ineq}$ impurity models
% leads to good results, with fair control of spatial numerical noise.

@@

In panel (b) of \figu{fig:KMflake} we report the spatially resolved
interaction dependency of $I(\rightarrow:\leftarrow)$, as a direct 
observation of the enhanced local correlations at the boundary of the
finite system. While it is usually assumed that edge sites must be
more correlated than internal ("bulk") sites, based on the reduced
lattice coordination, the inter-spin mutual information quantifies directly the phenomenon: the edge correlations,
marked in lighter green, are decidedly higher as long as the system 
remains in a weakly magnetized regime. When the local magnetization
exceeds half of its saturation value, all local correlations start to decrease,
as spin fluctuations are damped by ordering. Significantly, the edge 
correlations get lower than the bulk correlations as the maximum is 
crossed, consistently with the observation that their magnetization
is always higher and so their fluctuations are frozen faster.
Finally, we observe that the nanosystem of \figu{fig:KMflake} has 
finite local magnetization for arbitrary low interaction strengths,
in striking contrast with the infinite lattice. This is a well-known
effect of quantum confinement, as proposed in early works on graphene
flakes \cite{Valli2016PRB,Valli2018NL}.

>>>>

The code implementation follows the same guidelines discussed above for the other examples. The user is required to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone. 
This is employed to evaluate the local interacting Green's function:
$$
G_\mathrm{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z)
\right]^{-1}, 
$$
entering in self-consistency conditions $\GG^{-1}=G_\mathrm{loc}^{-1}+\Sigma$. 
Here  $\z\in\CCC$, $N_k$ is the number of $k$-points   and $\Sigma(z)$ is the self-energy matrix  obtained from the DMFT solution of the problem.  
Moreover, the $H(k)$ is used to  construct the {\it renormalized} topological Hamiltonian \cite{Wang2010PRL,Wang2012PRX,Blason2023PRB}
$$
H_\mathrm{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant matrix. The matrix $H_\mathrm{top}$ describes the
low-energy properties of the many-body solution and can be used to characterize the topological nature of the interacting system  \cite{Gurarie2011PRB,Wang2012PRX,Wagner2023NC,Blason2023PRB,Bau2024PRB}.

@@

To complete the overview of \NAME features, we consider another
model of interacting quantum spin-Hall insulator defined on the honeycomb lattice, with two inequivalent sites in the unit cell. Neglecting the Rashba spin-orbit interaction and discarding any ionic character of the unit cell, the noninteracting Kane-Mele Hamiltonian \cite{Kane2005PRLa,Kane2005PRL} can be written in momentum space as
\begin{align*}
    H_\mathrm{KM} = \sum_{{k}} \psi^\dagger_{{k}} H({k}) \psi_{{k}},
\end{align*}
with $\psi_{{k}} = [c_{{k},\mathrm{A},\up},\,c_{{k},\mathrm{B},\up},\,c_{{k},\mathrm{A},\dw},\,c_{{k},\mathrm{B},\dw}]$ and
    \begin{align}
    H({k}) 
    &= x_{{k}}\Gamma_{01} + y_{{k}}\Gamma_{02} + \delta_{{k}}\Gamma_{33}, \label{eq:Hk_kanemele}\\[1mm]
   x_{{k}}&=t\bigl[\cos({k}\cdot{u})+\cos({k}\cdot{v})+1\bigr],\nonumber\\
   y_{{k}}&=t\bigl[\sin({k}\cdot{u})+\sin({k}\cdot{v})\bigr],\nonumber \\
   \delta_{{k}}&=2\lambda_\mathrm{so}\bigl[\sin({k}\cdot{u})-\sin({k}\cdot{v}) - \sin({k}\cdot{u} -{k}\cdot{v})\bigr], \nonumber
\end{align}
where ${u} = (3a/2, \sqrt{3}a/2)$, and ${v} = (3a/2, -\sqrt{3}a/2)$ 
are suitable basis vectors for the honeycomb lattice, $t$ is the
nearest-neighbor hopping amplitude and $\lambda_\mathrm{so}$ is 
the amplitude of a complex next-nearest-neighbor hopping,
with a spin-dependent $\pm\tfrac{\pi}{2}$ chiral phase arising from spin-orbit coupling, which is well-known to open a topological gap at the Fermi level \cite{Kane2005PRLa,Kane2005PRL}.
The spinorial $4\times4$ matrices $\Gamma_{a\ell} = \sigma_a\otimes\tau_\ell$ are 
defined in terms of Pauli matrices $\sigma_a$, $\tau_\ell$ referred, respectively 
to spin and sublattice degrees of freedom.
% , with the convention that 
% $\sigma_0$ and $\tau_0$ are $2\times2$ identity matrices.
To investigate the effects of electron-electron repulsion we consider a local Hubbard term \cite{Hohenadler2013ROPIP,Rachel2018ROPIP}:
\begin{equation}
    H_\mathrm{KMH}  = H_\mathrm{KM} 
    + U\sum_{i}\Biggl[
                        \left(n_{i,\up} - \frac{1}{2}\right)
                        \left(n_{i,\dw} - \frac{1}{2}\right)
                      \Biggr],
    \label{eq:KMHmodel}
\end{equation}
where $n_{i,\sigma}=c^\dagger_{i,\sigma}c_{i,\sigma}$ are the 
local spin-density operators. The interaction is written in an explicitly 
particle-hole symmetric form so that the model is at half-filling for zero
chemical potential.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/flakes_edipack.pdf}\\[1mm]
    (c) \hspace{10cm} (d) \\
    \hspace{1.2cm}
    \includegraphics[width=0.33\linewidth]{figures/3NflakeU4.5_x.pdf} \hspace{1.2cm}
    \includegraphics[width=0.33\linewidth]{figures/3NflakeU6_x.pdf} \vspace{3mm}
    \caption{Kane-Mele-Hubbard model at the nanoscale.
    In (a) the local in-plane
    magnetization $\ibra S_x \iket = \ibra n_{i,\rightarrow}-n_{i,\leftarrow}\iket$. In
    (b) the local correlation as measured by the mutual 
    information between local in-plane spin eigenstates
    $I(\rightarrow:\leftarrow)$ \cite{BellomiaPhD,BellomiaKMH,Bellomia_intracorr}. 
    Both the (a) and (b) panels mark the edge and ``bulk'' 
    sites in lighter and darker colors, respectively.
    In (c) and (d) a visual representation of the given 
    nano-system at different values of $U/t$, as marked 
    respectively by the orange and red vertical lines 
    in panels (a) and (b). The area of the 
    circles is proportional to $I(\rightarrow:\leftarrow)$, 
    the length of the arrows to $\ibra S_x \iket$.
    All data are taken at $\lambda_\mathrm{so}=0.3t$ and $T=0$.}
    \label{fig:KMflake}
\end{figure}

\subsubsection{Kane-Mele-Hubbard model at the nanoscale}
The capabilities of the EDIpack2ineq layer go well beyond the
simple case of a two-site unit cell. Building on the previous example, we can readily adapt the calculation to deal with a nanoscopic~\cite{Amaricci2014PRA} system with $N_\mathrm{ineq}$ inequivalent atoms, such as hexagonal flakes
\cite{Valli2016PRB,Valli2018NL}.
We consider a 
$N_\mathrm{ineq}\times N_\mathrm{ineq}$ local Hamiltonian 
$H_\mathrm{loc}$ with open boundary conditions, and implementing
the self-consistent equations with a trivial single momentum 
$k=0$. In this configuration the system develops an inhomogeneous magnetization, so that in general no particular lattice symmetry group, e.g. N\'eel, can be exploited. 
% Some lowered point symmetry can be found, but a 
% safer approach of solving all the $N_\mathrm{ineq}$ impurity models
% leads to good results, with fair control of spatial numerical noise.

@@

In panel (b) of \figu{fig:KMflake} we report the spatially resolved
interaction dependency of $I(\rightarrow:\leftarrow)$, as a direct 
observation of the enhanced local correlations at the boundary of the
finite system. While it is usually assumed that edge sites must be
more correlated than internal (``bulk'') sites, based on the reduced
lattice coordination, the inter-spin mutual information quantifies directly the phenomenon: the edge correlations,
marked in lighter green, are decidedly higher as long as the system 
remains in a weakly magnetized regime. When the local magnetization
exceeds half of its saturation value, all local correlations start to decrease,
as spin fluctuations are damped by ordering. Significantly, the edge 
correlations get lower than the bulk correlations as the maximum is 
crossed, consistently with the observation that their magnetization
is always higher and so their fluctuations are frozen faster.
Finally, we observe that the nanosystem of \figu{fig:KMflake} has 
finite local magnetization for arbitrary low interaction strengths,
in striking contrast with the infinite lattice. This is a well-known
effect of quantum confinement, as proposed in early works on graphene
flakes \cite{Valli2016PRB,Valli2018NL}.
